RNApyro

### 2012-10-04 Recomb/methods_RECOMB.tex
fixed probs in tex
formalized EI

\subsection{Energy Model}
The energy we will be composed of two function, $\text{ES}^{\beta}_{ab\to a'b'}$ and 
$\text{EI}^{\Omega}_{(i,j),ab}$. The former is equal to the 
stacking energy of the base pair with nucleotides $ab$ on top of the base pair with nucleotides 
$a'b'$, as set in the NNDB~\cite{Turner2010}. If one of the base pair is not valid (i.e. not in 
$\{\text{GU},\text{UG},\text{CG},\text{GC}, \text{AU or UA}\}$, the value is a parameter 
$\beta \in [1,\infty]$. This allows
to completely forbid a sequence where a base pair is non valid, when $\beta = \infty$ or only 
penalize it.
$\text{EI}^{\Omega}_{(i,j), a'b'}$ is the average of the sum of differences between the isostericity
of base pairs at positions $(i,j)$ in $\Omega$ and $s_is_j$, and the isostericity of base pairs
 at positions $(i,j)$ in $\Omega$ and $ab$. If gives us an indication 
 if the base pair $ab$ is more isosteric to the set $\Omega$ than the one on the sequence 
 $\omega$. Where the isostericity values 
 between any pair of canonical base pairs is  as defined in~\cite{Stombaugh2009}. 
 Let be $\alpha\in[0,1]$, it 
 will be used to balance the weight given to the stacking energy or the isostericity.

@@

We are now interested in knowing, under our model, 
 the probability that a given position is a given nucleotide.
We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of mutations allowed. 
$$
	\mathbb{P}(s_i = x\mid s,\Omega, S,M):=\left\{
	\begin{array}{ll}
		\displaystyle
		\frac{
			\displaystyle
			\sum_{m=0}^{M}
			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
		}{
			\displaystyle
			\sum_{m=0}^{M}
			\Z{0,n-1}{m}{X,X}
		}
		&\text{If }S_i = -1\\
		\displaystyle
		\frac{
			\displaystyle
			\sum_{m=0}^{M}
			\sum_{\substack{b\in Bases\\\Kron_{xb,s_is_k\leq m}}}
			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
			\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
			\Z{i+1,k-1}{m'}{x,b}
		}{
			\displaystyle
			\sum_{m=0}^{M}
			\Z{0,n-1}{m}{X,X}
		}
		&\text{If }S_i=k>i\\
		\displaystyle
 		\frac{
			\displaystyle
			\sum_{m=0}^{M}
			\sum_{\substack{b\in Bases\\\Kron_{bx,s_ks_i\leq m}}}
			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
			\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
			\Z{k+1,i-1}{m'}{b,x}
		}{
			\displaystyle
			\sum_{m=0}^{M}
			\Z{0,n-1}{m}{X,X}
		}
		&\text{If }S_i=k<i
	\end{array}\right.
$$

In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
\begin{description}
\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, we are concerned by the weights
over all sequences which have at position $i$ nucleotide $x$, which is exactly the sum
of the values of $\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}$, for all $m$ between $0$ and $M$.
\item[$S_i=k>i$:] Since we need to respect the derivation of the secondary structure $S$, if 
position $i$ is paired, we must consider the two partition functions. The \emph{outside} of the 
base pair, and the \emph{inside}, for all possible values for the nucleotide at position $k$, and
all possible distribution of the mutant positions between the inside and outside of the base pair.
\item[$S_i=k<i$:] Same as above, but with $i$ pairing with a lower position.
\end{description}

>>>>

\subsection{Energy Model}
The energy we will be composed of two function, $\text{ES}^{\beta}_{ab\to a'b'}$ and 
$\text{EI}^{\Omega}_{(i,j),ab}$. The former is equal to the 
stacking energy of the base pair with nucleotides $ab$ on top of the base pair with nucleotides 
$a'b'$, as set in the NNDB~\cite{Turner2010}. If one of the base pair is not valid (i.e. not in 
$\{\text{GU},\text{UG},\text{CG},\text{GC}, \text{AU or UA}\}$, the value is a parameter 
$\beta \in [1,\infty]$. This allows
to completely forbid a sequence where a base pair is non valid, when $\beta = \infty$ or only 
penalize it.
$\text{EI}^{\Omega}_{(i,j), a'b'}$ is the average of the sum of differences between the isostericity
of base pairs at positions $(i,j)$ in $\Omega$ and $s_is_j$, and the isostericity of base pairs
 at positions $(i,j)$ in $\Omega$ and $ab$. If gives us an indication 
 if the base pair $ab$ is more isosteric to the set $\Omega$ than the one on the sequence 
 $\omega$. Formally, given $s$, $\Omega$, two positions $(i,S_i)$ and two nucleotides $a,b\in B$:
 $$
 	\text{EI}^{\Omega}_{(i,j),ab}:=
	\frac{
		\displaystyle
		\sum_{s'\in\Omega}
		\left(
			\text{ISO}((s_i',s_j'),(a,b))
		\right)-
		\left(
			\text{ISO}((s_i',s_j'),(s_i,s_j))		
		\right)
	}{
		\displaystyle
		|\Omega|
	}
 $$
  Where $\text{ISO}((a',b'),(a,b))$ is the isostericity values 
 between a the canonical base pairs $(a',b')$ and $(a,b)$  as defined in~\cite{Stombaugh2009}. 
 Let be $\alpha\in[0,1]$, it 
 will be used to balance the weight given to the stacking energy or the isostericity.

@@

We are now interested in knowing, under our model, 
 the probability that a given position is a given nucleotide.
We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of mutations allowed. 
$$
	\mathbb{P}(s_i = x\mid s,\Omega, S,M):=\left\{
	\begin{array}{ll}
		\displaystyle
		\frac{
			\displaystyle
			\sum_{m=0}^{M}
			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
		}{
			\displaystyle
			\sum_{m=0}^{M}
			\Z{0,n-1}{m}{X,X}
		}
		&\text{If }S_i = -1\\
		\displaystyle
		\frac{
			\displaystyle
			\sum_{m=0}^{M}
			\sum_{\substack{b\in Bases\\\Kron_{xb,s_is_k\leq m}}}
			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
     	 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(i,k),xb}}{RT}}
			\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
			\Z{i+1,k-1}{m'}{x,b}
		}{
			\displaystyle
			\sum_{m=0}^{M}
			\Z{0,n-1}{m}{X,X}
		}
		&\text{If }S_i=k>i\\
		\displaystyle
 		\frac{
			\displaystyle
			\sum_{m=0}^{M}
			\sum_{\substack{b\in Bases\\\Kron_{bx,s_ks_i\leq m}}}
			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
     	 e^{\frac{-(1-\alpha)\text{EI}^{\Omega}_{(k,i),bx}}{RT}}
			\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
			\Z{k+1,i-1}{m'}{b,x}
		}{
			\displaystyle
			\sum_{m=0}^{M}
			\Z{0,n-1}{m}{X,X}
		}
		&\text{If }S_i=k<i
	\end{array}\right.
$$

In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
\begin{description}
\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, we are concerned by the weights
over all sequences which have at position $i$ nucleotide $x$, which is exactly the sum
of the values of $\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}$, for all $m$ between $0$ and $M$.
\item[$S_i=k>i$:] Since we need to respect the derivation of the secondary structure $S$, if 
position $i$ is paired, we must consider the two partition functions. The \emph{outside} of the 
base pair, and the \emph{inside}, for all possible values for the nucleotide at position $k$, and
all possible distribution of the mutant positions between the inside and outside of the base pair. We also add the term of isostericity for this specific base pair.
\item[$S_i=k<i$:] Same as above, but with $i$ pairing with a lower position.
\end{description}
